### 2020-05-03 manuscript.tex
Imrove Step by step circuit explanation #16

\documentclass[pra,showkeys,twocolumn,showpacs]{revtex4-1}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{array}
\usepackage{color}
\usepackage{soul}
%\usepackage{floatflt}
\usepackage{graphicx}
\DeclareMathOperator{\tr}{Tr}
\newcommand*{\eq}[1]{(\ref{#1})}
\newcommand*{\eqs}[2]{(\ref{#1})--(\ref{#2})}
\newcommand{\exi}{\tr{\{e^{-\frac\beta2\sum_k\lambda_k\gamma_k^\dag\gamma_k}\}}}
\renewcommand{\l}{\left(}
\renewcommand{\r}{\right)}

\usepackage[inline]{enumitem} % inline enumeration

\begin{document}
\title{Quantum-assisted unsupervised data clustering}

\author{I.~D.~Lazarev} 
\affiliation{Institute of Problems of Chemical Physics of Russian Academy of Sciences, Acad. Semenov av. 1, Chernogolovka, Moscow Region, Russia, 142432}
\affiliation{ Faculty of Fundamental Physical-Chemical Engineering, Lomonosov Moscow State University, GSP-1, Moscow, Russia 119991}

\author{Marek Narozniak}
\affiliation{New York University Shanghai, 1555 Century Ave, Pudong, Shanghai 200122, China}
\affiliation{Department of Physics, New York University, New York, NY, 10003, USA.}

\author{Tim Byrnes}
\affiliation{New York University Shanghai, 1555 Century Ave, Pudong, Shanghai 200122, China}
\affiliation{State Key Laboratory of Precision Spectroscopy, School of Physical and Material Sciences, East China Normal University, Shanghai 200062, China}
\affiliation{NYU-ECNU Institute of Physics at NYU Shanghai, 3663 Zhongshan Road North, Shanghai 200062, China}
\affiliation{National Institute of Informatics, 2-1-2 Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan}
\affiliation{Department of Physics, New York University, New York, NY 10003, USA}

\author{A.~N.~Pyrkov} 
\email{Email address:pyrkov@icp.ac.ru}
\affiliation{Institute of Problems of Chemical Physics of Russian Academy of Sciences, Acad. Semenov av. 1, Chernogolovka, Moscow Region, Russia, 142432}

\date{\today}

\begin{abstract}
	Unsupervised machine learning is one of the main techniques in artificial intelligence. 
	Quantum computers offer opportunities to speed up such machine learning techniques. 
	Here, we propose a realization of quantum assisted unsupervised data clustering using the self-organizing feature map artificial neural networks. 
	We make a proof-of-concept realization of one of the central components of the approach on the IBM Q Experience 
	and show that it allows us to reduce number of calculations in a number of clusters. 
	We compare the results with the classical algorithm on a toy example of unsupervised text clustering.  
\end{abstract}

%\pacs{}
\maketitle

\section{Introduction}
The combination of big data and artificial intelligence ---  dubbed the fourth industrial revolution --- has profoundly affected the modern economy in a plethora of different ways from robotics to agriculture \cite{Lecun2015, ghahramani2015,schwab2017,esteva2019, tyrsa2017}. 
Contemporary artificial intelligence methods based on neural networks also have the potential to enhance the role of novel analytical methods in science and engineering \cite{kaggle2014, radovic2018, butler2018, radovic2018}. 
Paradoxically, the exact mechanism of why neural networks are so powerful remain unknown (in many cases it is regarded as a black box).  
It has been speculated that limits to the neural network approach based on the computational power of von Neumann architecture are being approached, 
and improvements appear only due to heuristic escalation of complexity \cite{marcus2018,sze2017,kourtis2020}.

In particular, the self-organizing feature map (SOFM) \cite{kohonen1990,kohonen1996,kohonen1997}, is a type of artificial neural network (ANN) 
that is trained in an unsupervised manner. 
SOFMs are used in many areas \cite{vilibic2016, guido1998, doszkocs1990, jones2012,mori2019,corsello2017,zhu2018,chea2016} 
and in comparison with many other artificial neural networks they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}. 
The SOFMs with a small number of nodes are similar to the K-means algorithm 
but for larger SOFMs they represent data in a fundamentally topological way that allows one to do dimensionality reduction.  
Once it is trained, the map can classify a vector from the input space by finding the node with the smallest distance metric.

Meanwhile, there has been much interest recently in applying quantum computing techniques to machine learning \cite{dunjko2018, biamonte2017, schuld2014, carleo2019}. 
The main focus of early works in quantum machine learning (QML) was on the use of quantum computers to perform basic linear algebra subroutines with a quantum speedup (such as Fourier transform and solving systems of linear equations \cite{wiebe2012,harrow2009,childs2017}) by utilizing quantum phenomena such as quantum entanglement and quantum superposition \cite{biamonte2017, schuld2014}.  
Quantum versions were developed for linear regression, principal component analysis, support vector machine, K-means algorithm and others \cite{lloyd2013,lloyd2014,dunjko2016,paparo2014,rebentrost2014}. 
More recently, there has been attention on developing quantum neural networks \cite{kamruzzaman2019, schuld2014b, jeswal2019, broughton2020}. 
The interest in quantum neural networks was inspired by progress in experimental quantum computing when it became possible to use parametrized quantum circuits, 
where the parameters behave much like the weights of a neural network \cite{lewenstein1994}. 
In particular, quantum algorithms for training and evaluating feedforward neural networks was developed, 
which are one of the most usable neural network models \cite{allcock2018, tacchino2019}. 
Recent results connecting the classical Bayesian approach to deep learning allowed for the development of a new algorithm for Bayesian deep learning on quantum computers \cite{zhao2019}. 
The quantum models for convolutional neural networks, which may be suitable for the problems of learning of quantum states were also proposed \cite{cong2019, liu2019}. 
Quantum classification, tested on the MNIST dataset, via slow feature analysis based on the use of quantum Frobenius distance was proposed \cite{kerenidis2018}. 
Sublinear quantum algorithms for training linear and kernel-based classifiers were designed \cite{li2019}. 
Furthermore, continuous-variable quantum neural networks \cite{killoran2019}, 
quantum autoencoders \cite{bondarenko2019}, 
quantum approaches for reinforcement learning \cite{dunjko2017, nautrup2019, foesel2018} 
and some other \cite{rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, pyrkov2019, vinci2019, lu2019} were developed. 
At the same time, non-neural-network based hybrid quantum classical algorithms have become a new direction of significant interest \cite{mcclean2016,arute2020,akshay2020}. 
Such hybrid algorithms involve as a part some quantum circuits for speed-up and usually trained in a standard classical learning loop. 
In particular, the Quantum Approximate Optimization Algorithm was developed to find approximate solutions to combinatorial optimization problems \cite{farhi2014,farhi2016} 
and designed to problems such as MAX-CUT, Grover’s search algorithm and so on \cite{arute2020,akshay2020,wang2018,jiang2017,huang2019,wecker2016,pagano2019}. 
Another example of well-known hybrid quantum classical algorithms is Variational Quantum Eigensolver (VQE) for applications in quantum chemistry 
\cite{kandala2017,aspuru-guzik2005,lanyon2010,peruzzo2014}.
Currently, it is believed that implementation of quantum neural networks and hybrid quantum classical algorithms can be the main test bed to achieve practical quantum supremacy on noisy intermediate scale quantum (NISQ) devices.

In this paper, we analyze  a hybrid quantum assisted SOFM (QASOFM) 
and apply it to the data clustering problem in an unsupervised manner on a toy example of clustering paper abstracts.   
We implement the quantum assisted SOFM in such a way 
that it becomes possible to use the Hamming distance as a distance metric for the training 
and to reduce the number of distance calculations in the number of clusters. 
We propose an optimized circuit for realizing the Hamming distance on a quantum machine. 
We then apply it to a toy example of data for clustering paper abstracts 
and give a proof-of-concept realization of the quantum assisted SOFM on the IBM Q experience quantum computer \cite{ibmq} 
and compare it to the classical case.

\begin{figure}
  	\includegraphics[width=0.95\columnwidth]{sofm_fitting.png}
	\caption{
		Schematic illustration of the clustering problem considered in this paper.  
		Blue dots represent clusters and red dots are data points. 
		The training process moves clusters to fit the data points. 
		Note that there are fewer clusters than data points, 
		which is the essence of dimensionality reduction, 
		and is what permits the model to generalize the data.
	}
	\label{fig:sofm_fitting}
\end{figure}

\section{The quantum assisted self-organizing feature map (QASOFM)}
\label{sec:qasofm}

\subsection{The classical algorithm}

The SOFM is one of the most widely-used unsupervised learning methods used in various areas of modern science. 
It was first proposed by Kohonen as a self-organizing unsupervised learning algorithm which produces feature maps similar to those occurring in the brain \cite{solan2001}. 
The SOFM algorithm operates with a set of input objects, each represented by a $N$-dimensional vector, 
and describes a mapping from a higher-dimensional input space to a lower-dimensional map space, commonly a bi-dimensional map.

The input dimensions are associated with the features, 
and the nodes in the grid (called cluster vectors) are assigned the $N$-dimensional vectors. 
The components of these vectors are usually called weights. 
Initially the weight components are chosen randomly 
and topological distances between neurons given. 
We then can train our SOFM adjusting the components through the learning process which occur in the two basic procedures of 
selecting a winning cluster vector, also called the best matching unit (BMU), and updating its weights (Fig.~\ref{fig:sofm_fitting}). 
More specifically, they consist of four step process: 
\begin{enumerate*}
\item selecting an input vector randomly from the set of all input vectors; 
\item finding a cluster vector which is closest to the input vector (BMU); 
\item adjusting the weights of the BMU and neurons close to it on feature map in such a way 
	that these vectors  becomes even closer to the input vector; 
\item repeating this process for many iterations until it converges.
\end{enumerate*}

On a step $t$ when the BMU $w_{c}$ is selected, 
the weights of the BMU and its neigbours on feature map are adjusted according to
%
\begin{equation}
    \label{eq:learning}
	\vec{w_{i}}(t + 1) 
	= \vec{w_i}(t)
	+ \theta(c, i, t) \alpha(t) 
		\left(\vec{x}(t) - \vec{w_i}(t)\right),
\end{equation}
%
where  $\alpha(t)$ is the learning rate and $\theta(c, i, t)$ is the neighbor function,
which defines the topological neighbor neurons to be updated. 
Note that neighbor function depend on distance on feature map given initialy, 
but not distance metrics between vectors.

Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors in space one at a time in a way 
that ensures each move is following the current trends inferred from their distances to the input objects. 
A visualisation of this process is shown in Fig. \ref{fig:sofm_fitting}.

In SOFM original version was designed to cluster real-valued data
and the winning cluster vector is selected based on the Euclidean distance between an input vector and the cluster vectors.
The paper deals with binary vectors clustering problem 
and for binary data the Hamming distance is more suitable \cite{appiah2009, santana2017}.
Using a known technique of encoding classical information into a quantum register, 
based on probabilistic quantum memories \cite{trugenberger2001},
we introduce an optimized algorithm for calculating the matrix of Hamming distances 
between each pair of binary vectors  of two sets taking advantage of quantum parallelism.

\subsection{Optimized quantum scheme for Hamming distance calculation}
\label{subsec:qcircuit}

%%%%%
\begin{figure}[t]
	\includegraphics[width=\columnwidth]{qcircuit.eps}
	\caption{
		Quantum circuit for the quantum parallelized Hamming distance calculating  between all pairs of binary vectors from two sets ${X}$ and ${Y}$ encoded \cite{trugenberger2001} in $X$ and $Y$ quantum registers   respectively.
		First, we encoded information about pairwise different qubits in a quantum state of the $X$-register with applying the CNOT gates. 
		Second, Hamming distance values are extracted into the amplitudes of superposition with the controled rotation around $z$-axis gate~(\ref{eq:controled_rotation}) and Hadamard gates. 
		Finally, a quantum state of the $X$-register returned to the initial basis for information retrieval. 	
	} 
	\label{fig:qcircuit}
\end{figure}
%%%%%

%%%%%
\begin{figure}[t]
	\includegraphics[width=0.95\columnwidth]{distance_matrix.png}
	\caption{
		(a) Hamming distance matrix between two data sets of binary vectors. 
		Values of distance that are less than the median distance value are marked black. 
		Classical simulation of the quantum circuit shows perfect agreement to theoretical calculations. 
		(b) Result obtained on the IBM Q Experience ``ibmq\_16\_melbourne'' backend. 
		We can see good agreement between the distance matrices calculated classically and on the IBM Q Experience.  
	}
	\label{fig:distance_matrix}
\end{figure}
%%%%%

The overall procedure involves two registers of $n$ qubits each, denoted $\left| X \right\rangle$ and $\left| Y \right\rangle$, along with a single auxiliary qubit $\left| a \right\rangle$. 
During the whole process, the $\left| Y \right\rangle$ register is used to store the cluster states.  
At the beginning and end of the procedure the $\left| X \right\rangle$ register stores the input vectors.  
Meanwhile, during the procedure it stores the differences between input vectors and cluster states.

Let us assume we have $k$ input vectors and $l$ cluster states. 
The $i$th input vector and $j$th cluster vector are respectively denoted as $\left| x_i \right\rangle$, $\left| y_j \right\rangle$. 
The registers $\left| X \right\rangle$ and $\left| Y \right\rangle$ are initialized to store the input vectors and cluster vectors according to
%
\begin{align}
    \left| X \right\rangle  & = \frac{1}{\sqrt{k}} \sum\limits_{i=1}^{k} \left| x_i \right\rangle,  \\
    \left| Y \right\rangle&  = \frac{1}{\sqrt{l}} \sum\limits_{j=1}^{l} \left| y_j \right\rangle .
    \label{eq:encodnig}
\end{align}
% 
The two registers along with the auxiliary qubit comprise the initial state of the quantum computer according to
%
\begin{equation} 
\left| \psi_0 \right\rangle = 
    \left| X \right\rangle
    \left| Y \right\rangle 
    \left| a \right\rangle
    \label{eq:initial_state}
\end{equation}
%
where $\left| a \right\rangle$ is an auxiliary qubit in the state $\left| 0 \right\rangle$ initially.

Given this initial state we may begin the processing of the problem. We start by applying a CNOT gate on $\left| x^{(\alpha)} y^{(\alpha)} \right\rangle$ and $\alpha = 1 \dots n$

\begin{equation}
    | \psi_1 \rangle  =  
    \frac{1}{\sqrt{kl}} \sum_{i, j=1}^{k} 
    | d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \rangle 
    | y^{(1)}_j, \dots, y^{(n)}_j \rangle
    | 0 \rangle 
\end{equation}
%
where $d^{(\alpha)}_{ij} = \mathrm{CNOT}(y^{(\alpha)}_i, x^{(\alpha)}_j)$, and $\alpha = 1 \dots n$  is the qubit index in the register. 
At this stage of the computation the $\left| X \right\rangle$ no longer stores the input vectors,
instead it stores the information about pairwise different qubits between the input vector $\{X\}$ and cluster vector $\{Y\}$. 
Next, for each pair $\{X\}$ and $\{Y\}$, the accumulated information of all the differences is projected onto the amplitude of the superposed state. 
This is achieved by applying the Hadamard gate on auxiliary qubit, 
followed by a controlled rotation around $z$-axis gate on $\left| Xa \right\rangle$ defined as
%
\begin{equation}
    \label{eq:controled_rotation}
    C_{R_z}(\phi) = 
    \begin{pmatrix}
        1 & 0 & 0 & 0 \\
        0 & e^{-i \frac \phi 2} & 0 & 0 \\
        0 & 0 & 1 & 0 \\
		0 & 0 & 0 & e^{i \frac \phi 2}
    \end{pmatrix} .
    \quad \phi = \frac{\pi}{n}
\end{equation}
%
Finally, another Hadamard gate is applied on the ancilla qubit (see Fig. \ref{fig:qcircuit}).

After the first Hadamard on the ancilla qubit the state is
%
\begin{equation}
    \left| \psi_2 \right\rangle = 
    \frac{1}{\sqrt{kl}}\sum\limits_{i, j=1}^{k} 
    \left| d_{ij} \right\rangle 
    \left| y_j \right\rangle
    \dfrac{(\left| 0 \right\rangle + \left| 1 \right\rangle)}{\sqrt{2}}  .
\end{equation}
%
Applying the controlled rotation around $z$-axis gate on $\left| x^{(\alpha)} a \right\rangle$ where $\alpha = 1\dots n$ and the state then becomes
%
\begin{multline}
    \left| \psi_3 \right\rangle  
	= \dfrac{1}{\sqrt{2kl}} \sum\limits_{i, j=1}^{k} 
		\exp\left(
		    \dfrac{-i \pi}{2n}
		    \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij}
		\right)
		\left| d_{ij} \right\rangle 
		\left| y_j \right\rangle 
		\left| 0 \right\rangle \\
    + \dfrac{1}{\sqrt{2kl}} \sum\limits_{i, j=1}^{k}
		\exp\left(
		    \dfrac{i \pi}{2n}
		    \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij} 
		\right)
		\left| d_{ij} \right\rangle  
		\left| y_j \right\rangle 
		\left| 1 \right\rangle
\end{multline}
%
Applying another Hadamard on the ancilla qubit we obtain
%
\begin{multline}
    \left| \psi_4 \right\rangle = 
    \frac{1}{\sqrt{kl}}\sum\limits_{i, j=1}^{k} 
        \left[ 
			\cos\left(
			    \dfrac{\pi}{2n}
			    \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij}
			\right)
			\left| d_{ij} \right\rangle 
			\left| y_j \right\rangle 
			\left| 0 \right\rangle\right.
			\\+ 
			\left. i \sin\left(
			    \dfrac{\pi}{2n} 
			    \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij} 
			\right)
			\left| d_{ij} \right\rangle 
			\left| y_j \right\rangle 
			\left| 1 \right\rangle
		\right] .
\end{multline}
%
This completes the step for projecting differences between pairs of $\{X\}$ and $\{Y\}$ onto the amplitude of the auxiliary qubit. 
The process is done in the $x$-basis, achieved by the surrounding Hadamard gates. 
There are two possible measurement outcomes of the auxiliary qubit.  
Each pair of $\{X\}$ and $\{Y\}$ forms a subspace of the Hilbert space, 
the controlled rotation around $z$-axis gate ensures to change amplitudes of those outcomes within this subspace depending on how different the spin configurations between $\{X\}$ and $\{Y\}$ are.

At this stage, the information regarding the differences between pairs of $\{X\}$ and $\{Y\}$ \hl{encoded in the amplitudes, in order to extract the Hamming distances between the relevant $\left| x_i \right\rangle$, $\left| y_j \right\rangle$ we return to our initial basis} for register $\left| X \right\rangle$ by applying pairwise CNOT gates:
%
\begin{multline}
    \left| \psi_f \right\rangle =  
    \frac{1}{\sqrt{kl}}\sum\limits_{i, j=1}^{k} 
	\left[ 
	    \cos\left(
	        \dfrac{\pi}{2n}
	        \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij} 
	    \right)
        \left| x_i \right\rangle 
        \left| y_j \right\rangle 
        \left| 0 \right\rangle\right.
        \\+
        \left. i \sin\left(
            \dfrac{\pi}{2n}
            \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij} 
        \right)
        \left| x_i \right\rangle 
        \left| y_j \right\rangle 
        \left| 1 \right\rangle
    \right] .
\end{multline}
%
This makes $\{X\}$ store the input vectors again, as in the initial step 
and preserves the amplitudes of the auxiliary qubit which are proportional to how different each pairs of $\{X\}$ and $\{Y\}$ are.

We thus have the Hamming distances encoded into the amplitudes of the final state. 
From the statistics of the measurement outcomes, the distance matrix between two data sets of binary vectors can be obtained (Fig.~\ref{fig:distance_matrix}). 
In this case, the biggest amplitude of the measurement result coincides with the smallest Hamming distance when the measurement result of the ancilla qubit is 0. 
If the ancilla qubit is 1, the smallest amplitude of the measurement result coincides with the smallest Hamming distance.

Measuring the Hamming distance of a particular pair of input vectors $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$ consists of extracting the relevant amplitude from the subspace that those states form, 
this can be done using the following projection operator
%
\begin{align}
\Pi_{i,j} = &\left| x_i \rangle\langle x_i \right| \otimes \left| y_j \rangle\langle y_j \right| \otimes I .
\end{align} 
%
Using the above projection operator, the subspace of the Hilbert space formed by a particular pair of input and cluster vectors can be traced out as
%
\begin{align}
    \rho_{i,j} &= \text{Tr}_{1,\dots,2n} (\Pi_{i,j} \left| \psi_f \rangle\langle \psi_f \right| \Pi_{i,j}) .
\end{align}
%
From the state the following two amplitudes for the measurement results can be extracted
%
\begin{align}
    a_0(x_i,y_j) & = \frac{\left\langle 0 |\rho_{i,j}| 0 \right\rangle}{\text{Tr}(\rho_{i,j})}  \\
    a_1(x_i,y_j) & = \frac{\left\langle 1 |\rho_{i,j}| 1 \right\rangle}{\text{Tr}(\rho_{i,j})} .
\end{align}
%
In order to reduce noise we average the measurement results over different states of the ancilla qubit, 
thus the measured Hamming distance between the input vector $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$ is
%
\begin{align}
    d_{i,j}^H & \propto 1 - \frac{1}{2}(a_0(x_i,y_j) + (1-a_1(x_i,y_j))) .
\end{align}
%
The Hamming distance measured in this way is bounded $0 \leq d_{i,j}^H \leq 1$, 
where  $0$ indicates that $x_i$ are $y_j$ identical and $1$ means they are completely opposite in terms of their pairwise binary coordinates.

In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, 
our method allows one to reduce the number of gates in the circuit (at least in $2n$ one-qubit NOT gates)
and reconstruct all distance matrix in parallel and can be implemented with high fidelity.

%%%%%
\begin{figure}[t]
	\includegraphics[width=0.95\columnwidth]{vectorized_sample.png}
	\caption{
		(a) Representation of the data set of abstracts with the bag-of-words \cite{weikang2016} model is shown. 
		Each abstract is represented by a binary vector with 9 elements, corresponding to the 9 words on the horizontal axis. 
		The samples are sorted into groups (QML, MED, BIO) with 3 papers for each tag, for a total of 9 paper.    
		(b) The Hamming distance between each vectorized abstract is shown as a number in the matrix. 
	} 
	\label{fig:vectorized_sample}
\end{figure}
%%%%%

\section{Experimental demonstration of QASOFM}

We now show experimental results for a proof-of-concept demonstration of the algorithm introduced in the previous section, on a 16-qubit quantum computer provided by IBM Q Experience.  
We perform unsupervised data clustering for three sets of paper abstracts from three different fields: 
quantum physics \cite{qml0, qml1, qml2}, 
medicine \cite{med0, med1, med2} 
and biology \cite{bio0, bio1, bio2}. 
Each set consists of three random papers that focus on one of following topics: 
``Quantum Machine Learning'' (QML), 
``Cancer'' (MED) 
and ``Gene Expression'' (BIO). 
Abstracts were vectorized by the bag-of-words \cite{weikang2016} model in order to choose most defining words in each data set (see Fig.~\ref{fig:vectorized_sample}) \cite{mctear2016}.  
This model represents text as a multiset ``bag'' of its words taking into account only multiplicity of words. 
Preparing the bag-of-words we excluded the words that appear only in one abstract and more than in 4 abstracts and we also excluded the word ``level'' from consideration due to the frequent overlap between the clusters because it gives instabilities for both classical and quantum algorithms. 
We restricted our bag-of-word size to 9 of the most frequent words from the full bags-of-word  due to limitations of the number of qubits. 
After vectorizing and pre-processing  the data, the clusters are well separated with the Hamming distance.  
We observe that distances between the abstracts inside clusters are smaller than distances between the abstracts from different clusters, 
showing successful self-organization (Fig.~\ref{fig:vectorized_sample}).

In classical SOFM \cite{kohonen1990}, first a distance calculation between a a sample vector and all cluster vectors is made, 
then the closest cluster vector is shifted towards the sample vector. 
The complexity of algorithm, in the sense of the number of distance calculations, scales as $O(LMN)$, 
where $N$ is number of samples, 
$M$ is number of randomly sampled cluster vectors, 
and $L$ is number of the shifts of cluster vectors.  
In the QASOFM described in the previous section, distance calculations are realized on a quantum device (i.e. the IBM Q Experience) with the use of circuit presented in Fig. \ref{fig:qcircuit}. This approach allows one to reduce the number of operations in a number of cluster states with an optimized number of gates 
that are possible to realize on  quantum computing devices that are currently available. 
The circuit realized in such a manner 
that calculation of Hamming distance between the sample vector and all cluster vectors is realized in one operation. 
The complexity of the quantum assisted SOFM then scales as $O(LN)$.

In order to check that our algorithm gives the expected results we compare it to classical calculations of the distance matrix on two data sets of binary vectors, as shown in Fig. \ref{fig:distance_matrix}.  
We see good agreement between the distance matrices calculated classically and on the IBM Q Experience. 
The theoretical calculations and classical simulations show perfect agreement with each other (Fig. \ref{fig:distance_matrix}(a)). 
The difference between Figs. \ref{fig:distance_matrix} (a) and (b) appears due to noise in the currently available non fault-tolerant quantum processors.

An example of the QASOFM learning process is shown in Fig.~\ref{convergence}. 
Initially, the cluster vectors were randomly chosen (see Fig.~\ref{convergence}) 
and the label of sample distribution is shown for the zeroth epoch in Fig.~\ref{convergence}(a). In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix we use the standard initialization of QISKIT library. 
Each epoch of the algorithm consists of distance calculation between all data and cluster vectors 
and requires 9 distance calculations in the  quantum implementation ($N$ in general case) 
or 27 distance calculations for classical realization 
($MN$ in general case). 
After the distance calculation from each sample to all cluster vectors at each epoch we label each sample with the index of the closest cluster vector 
and shift the closest cluster vectors to the sample one. 
The shift is made by the change of the first binary element in the cluster vectors different from the sample one. 
The evolution of the labels presented on Fig. ~\ref{convergence}(c).  
Good convergence is already observed in the fourth epoch.

This proof of concept example shows that developing of noisy intermediate scale quantum hardware will allow to solve some practical problems in unsupervised manner with very simple encoding of categorical data to quantum register. 
Furthermore, it shows that it is possible to develop distance based hybrid quantum classical algorithms 
that speed up classical counterparts and can be implemented on near future quantum devices 
and that further developing distance based hybrid algorithms, 
that can learn in unsupervised manner, for categorical data such as genomic and so on has significant interest.

%%%%%
\begin{figure}[t]
	\includegraphics[width=0.95\columnwidth]{convergence.png}
	\caption{
		(a) Initial random binary vectors of cluster vectors (labeled as C0, C1, C2), 
		for the selected 9 words from the bag of words model. 
		(b) The result of applying our QASOFM implemented on the IBM Q Experience simulator backend. 
		Vectors mean cluster elements for BIO, MED, QML groups, 
		for the C0, C1, C2 cluster vectors, respectively. 
		(c) The evolution of label distribution on each learning epoch.
	} 
	\label{convergence}
\end{figure}
%%%%%

\section{Discussions}
We have developed a quantum assisted SOFM and showed a proof-of-concept experimental demonstration 
that it can be used to solve clustering problems in an unsupervised manner.  
The procedure of solving such clustering problem requires calculating the distance many times in iterative way, 
which is calculated using a hybrid quantum-classical procedure.  
We introduced an optimized circuit for Hamming distance calculations that can be implemented on currently available quantum computing devices with high fidelity. 
Our quantum circuit performs the distance-computing component of a classical SOFMs algorithm 
and in this way improve its performance. 
The complexity of the quantum assisted SOFM scales as $O(LN)$ 
while the complexity of the fully classical SOFM scales as $O(LMN)$, 
where $N$ is number of samples, $M$ is number of randomly sampled cluster vectors, 
and $L$ is number of the shifts of cluster vectors.
Due to wide use of classical SOFMs in different areas of modern research and technology, 
this can give opportunities for the use of QASOFM in practical applications in near term, outperforming classical algorithms.

\section*{Acknowledgments}
The work is supported by the RFBR-NSFC collaborative program (Grant No. 18-57-53007) and the State assignment (N. 0089-2019-0002). I.D.L. acknowledges support from the Russian Foundation for Basic Research, grant No.19-32-80004 and support from the Foundation for the Advancement of Theoretical Physics and Mathematics “BASIS” No.19-1-5-130-1. T.B. is supported by the Shanghai Research Challenge Fund; New York University Global Seed Grants for Collaborative Research; National Natural Science Foundation of China (61571301, D1210036A); the NSFC Research Fund for International Young Scientists (11650110425, 11850410426); NYU-ECNU Institute of Physics at NYU Shanghai; the Science and Technology Commission of Shanghai Municipality (17ZR1443600); the China Science and Technology Exchange Center (NGA-16-001); and the NSFC-RFBR Collaborative grant (81811530112).

%\section*{Author Contributions}

%I.D.L. and A.N.P. performed the calculations. All authors discussed the results and wrote the paper.

\section*{Conflict of interest}

The authors declare no conflict of interest.

\section*{Keywords}
quantum self-organizing feature map; quantum machine learning; neural network; quantum unsupervised data clustering; Hamming distance.

\bibliographystyle{naturemag}
\bibliography{bibliography}

%\begin{thebibliography}{99}
%\end{thebibliography}
\end{document}